\begin{thebibliography}{99}

\bibitem{Adams1979} Adams, D. (1979). \htmladdnormallink{\textit{The Hitchhiker's Guide to the Galaxy}}{http://www.amazon.com/dp/0345418913}. Pan Books, London (United Kingdom), 1979.

\bibitem{Aspect1982} Aspect, A.; et al. (1982). \textit{Experimental Realization of Einstein-Podolsky-Rosen-Bohm Gedankenexperiment: A New Violation of Bell's Inequalities}. Physical Review Letters 49: 91, American Physical Society, College Park (Maryland, United States), 1982.

\bibitem{Bassi2003} Bassi, A.; Ghirardi, G. (2003). \htmladdnormallink{\textit{Dynamical Reduction Models}}{http://arxiv.org/abs/quant-ph/0302164v2}. Physics Reports 379: 257, Elsevier, Amsterdam (The Netherlands), 2003.

\bibitem{CMS2011} CMS Collaboration (2011). \htmladdnormallink{\textit{Measurement of $W^+W^-$ Production and Search for the Higgs Boson in pp Collisions at $\sqrt s$ = 7 TeV}}{http://arxiv.org/abs/1102.5429v2}. Physics Letters B 699: 25, Elsevier, Amsterdam (The Netherlands), 2011.

\bibitem{Collins1984} Collins, J. (1984). \htmladdnormallink{\textit{Renormalization: an introduction to renormalization, the renormalization group and the operator product expansion}}{http://www.cambridge.org/9780521311779}. Cambridge University Press, Cambridge (United Kingdom), 1984.

\bibitem{Descartes1637} Descartes, R. (1637). \htmladdnormallink{\textit{Discours de la méthode pour bien conduire sa raison, et chercher la vérité dans les sciences}}{http://www.gutenberg.org/ebooks/13846}. Ian Maire, Leiden (The Netherlands), 1637. \htmladdnormallink{\textit{Discourse on the Method of Rightly Conducting One's Reason and of Seeking Truth in the Sciences}}{http://www.gutenberg.org/ebooks/59}. W. Blackwood and sons, London (United Kingdom), 1870.

\bibitem{Einstein1916} Einstein, A. (1916). \textit{Über die spezielle und die allgemeine Relativitätstheorie}. Vieweg, Braunschweig (Germany), 1917. \htmladdnormallink{\textit{Relativity: The Special and General Theory}}{http://www.gutenberg.org/ebooks/5001}. Methuen \& Co. Ltd., London (United Kingdom), 1920.

\bibitem{Fauvel2013} Fauvel, S. (2010). \htmladdnormallink{\textit{Quantum Ethics: A Spinozist Interpretation of Quantum Field Theory}}{http://quantum-ethics.org/}. CreateSpace Independent Publishing Platform, Scotts Valley (California, United States), 2013.

\bibitem{Feynman1964} Feynman, R. (1964). \htmladdnormallink{\textit{The Character of Physical Law: 6. Probability and uncertainty - the quantum mechanical view of nature}}{http://www.youtube.com/watch?v=aAgcqgDc-YM}. Messenger Lectures, Cornell University, Ithaca (New York, United States), 1964. The British Broadcasting Corporation, London (United Kingdom), 1965.

\bibitem{Freedman2001} Freedman, W. L.; et al. (2001). \htmladdnormallink{\textit{Final Results from the Hubble Space Telescope Key Project to Measure the Hubble Constant}}{http://arxiv.org/abs/astro-ph/0012376v1}. The Astrophysical Journal 553 (1): 47–72, IOP Publishing, Bristol (United Kingdom), 2001.

\bibitem{Friedman1922} Friedman, A. (1922). \textit{Über die Krümmung des Raumes}. Zeitschrift für Physik A 10 (1): 377–386, Springer, Berlin (Germany), 1922.

\bibitem{Hermanns1983} Hermanns, W. (1983). \textit{Einstein and the Poet: In Search of the Cosmic Man}. Branden Press, Boston (Massachusetts, United States), 1983.

\bibitem{Kogut1993} Kogut, A.; et al. (1993). \htmladdnormallink{\textit{Dipole Anisotropy in the COBE Differential Microwave Radiometers First-Year Sky Maps}}{http://arxiv.org/abs/astro-ph/9312056v1}. The Astrophysical Journal 419: 1–6, IOP Publishing, Bristol (United Kingdom), 1993.

\bibitem{Komatsu2010} Komatsu, E.; et al. (2010). \htmladdnormallink{\textit{Seven-Year Wilkinson Microwave Anisotropy Probe (WMAP) Observations: Cosmological Interpretation}}{http://arxiv.org/abs/1001.4538v2}. The Astrophysical Journal Supplement Series 192: 18, IOP Publishing, Bristol (United Kingdom), 2011.

\bibitem{Marin2009} Marin, J. M. (2009). \htmladdnormallink{\textit{‘Mysticism’ in quantum mechanics: the forgotten controversy}}{http://harvard.academia.edu/JuanMiguelMarin/Papers/204341}. European Journal of Physics 30: 807-822, IOP Publishing, Bristol (United Kingdom), 2009.

\bibitem{Neumann1932} Neumann, J. v. (1932). \textit{Mathematische Grundlagen der Quantenmechanik}. Springer, Berlin (Germany), 1932. \textit{Mathematical Foundations of Quantum Mechanics}. Princeton University Press, Princeton (New Jersey, United States), 1955.

\bibitem{Rabin1981} Rabin, J. M. (1981). \htmladdnormallink{\textit{Long Range Interactions In Lattice Field Theory}}{http://www.slac.stanford.edu/pubs/slacreports/slac-r-240.html}. Ph.D. Thesis SLAC-0240, Stanford (California, United States), 1981.

\bibitem{Spinoza1677} Spinoza, B. (1677). \textit{Ethica, Ordine Geometrico demonstrata}. Amsterdam (The Netherlands), 1677. \htmladdnormallink{\textit{The Ethics}}{http://www.gutenberg.org/ebooks/3800}. London (United Kingdom), 1883.

\end{thebibliography}
